\section{Problem Statement} \label{sec:problem statement}
This project is an implementation of first order logic, unification, and resolution through the setting of the Wumpus World. 
The Wumpus World is an explorable dungeon with the goal state of finding the room with gold in it; but there are pits, blocked rooms, and dangerous Wumpus monsters in the way. 
The agent has a bow and arrow at his disposal to kill the Wumpus and a refined inference engine to decide what rooms to go to and which actions to take.

The problem handled by this project focuses on the inference engine.
For dungeons of sizes 5 x 5 to 25 x 25, inference rules are used to always make the safest choice depending on information about the dungeon the agent has revealed.
The challenge is to decode how to state and use first-order clauses in a programmatic fashion. 

Evaluation will compare the logic-based agent to a simple, reactive agent.
The reactive agent always chooses what it believes is the most safe neighbor cell is based on only the current perception at time $t$ for room $C_{x,y}(t)$.
The metrics used to compare the two agents are:
\begin{enumerate}
	\item Number of decisions made.
    \item The gold was or was not found.
    \item Number of Wumpi killed by the agent.
    \item Number of times the agent fell in to a pit.
    \item Number of times the agent is killed by a Wumpus.
    \item Number of unique cells explored.
    \item The agent's score.\footnote{-1 point for each movement, -10 points for firing an arrow, +10 for killing a Wumpus, -1000 for dying, +1000 points for finding the gold.}
\end{enumerate}

\subsection{Hypothesis} \label{subsec:hypothesis}
Given the knowledge based agent (KB-agent) will have rules in place to never move into a potentially dangerous room until the inference engine can assure the room is safe, one can hypothesize the KB-agent will produce a better score than the reactive agent for every map size on average. 
For small map sizes\footnote{5x5 or 10x10} the reactive agent may outperform the KB-agent for minimizing number of decisions made and number of cells explored, but the reactive agent will be notably outperformed in all categories as the map size increases.  

Furthermore, The KB-agent will likely perform in NP time.
Run time will slow down dramatically as the search space becomes large due to exponential growth of logical comparisons that will be made\cite{Ansari}.
